---
title: "lab9_part2"
author: "Nick McManus"
date: "11/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(here)
library(broom)
```

## Read in data
```{r}
lizards <- read_csv(here("data", "lizards.csv"))
```

Get counts of lizards by species:
```{r}
lizards %>% 
  count(spp) %>% 
  arrange(-n)
```

## Make a subset

- UTST lizards
- Sites: IBPE (enclosure grassland), RABB (rabbit mesquite site), "EAST" (east tarbush site)
- tail condition is records (W = whole, B = broken)

```{r}
utst <- lizards %>% 
  filter(spp == "UTST") %>% 
  filter(site %in% c("IBPE", "RABB", "EAST")) %>% 
  filter(tail %in% c("B", "W"))
```

## Find counts and proportions of lizards with broken or whole tails at each site

```{r}
utst %>% 
  count(site, tail)
```

```{r}
tail_counts <- utst %>% 
  janitor::tabyl(site, tail)  # within janitor package

tail_proportions <- tail_counts %>% 
  janitor::adorn_percentages() %>%    # default is proportions by rows
  janitor::adorn_pct_formatting(digits = 2) %>%   #shows as percents with specific number of decimals
  janitor::adorn_ns()  # shows counts in ()
```

## Convert to a contingency table

Use   `column_to_rowname()`. 

```{r}
lizard_ct <- tail_counts %>% 
  column_to_rownames(var = "site")
```


## Chi-square

Are proportions of lizards that have broken or whole tails so different, we have enough evidence to reject the null hypothesis?

```{r}
lizard_x2 <- chisq.test(lizard_ct)
lizard_x2

lizard_tidy <- broom::tidy(lizard_x2)  # puts in tidyformat to make it easier to reference in in-line coding
```

There is a significant association (i.e. non-independence) between site and lizard tail breakage ($\chi$^2^(`r lizard_tidy$parameter`) = `r lizard_tidy$statistic`, *p* = `r lizard_tidy$p.value`).
























